---
title: "Using database"
author: ""
date: "23/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Connect to the database

- Now let's use the databases
- First connect to the database

```{r}
library(DBI)
library(tidyverse)

db <- dbConnect(RSQLite::SQLite(), "~/covid.sqlite")
```

## Run queries

### Get sample rows from each table 

- Get sample rows from each table
- Use `dbGetQuery()`
  - Run SQL queries there

```{r}
  dbGetQuery(db, "SELECT * FROM covid LIMIT 5")
  dbGetQuery(db, "SELECT * FROM demo LIMIT 5")
  dbGetQuery(db, "SELECT * FROM trump LIMIT 5")


```


### Bit more complex queries

1. Get the latest data for each county
2. Get the data on 1 March, 2021, and select county with more than 10 deaths

```{r}
dbGetQuery(db, "select * from covid where date = '2021-03-19'" )
dbGetQuery(db, "select * from covid where date = '2021-03-01' and deaths > 10" )

```
#### Ordering and aggregation

For the data above

1. Order counties with the highest number of death
2. Calculate the state level death figures


```{r}
dbGetQuery(db, "select * from covid where date = '2021-03-19' order by deaths desc" )
dbGetQuery(db, "select state, sum(deaths) from covid where date = '2021-03-19' group by state" )

```


### Merging tables

- merge covid and demo table by fips
- (advanced) calculate per 1000 cases and deaths

```{r}
## It's slow because the tables are not indexed. 
dbGetQuery(db, "select covid.*, demo.* from covid inner join demo on covid.fips = demo.fips")
dbGetQuery(db, "select covid.*, demo.*, covid.deaths / demo.totalPop * 1000 as deaths_per1000  from covid inner join demo on covid.fips = demo.fips")


```

## dblyr

Now we will use `dbplyr`

### Get the object for each table

```{r}
tb_covid <- tbl(db, "covid")
tb_demo <- tbl(db, "demo")

tb_covid
```


### Let's check the same thing we did:


#### Get sample rows from each table 

- Get sample rows from each table
- Use `dbGetQuery()`
  - Run SQL queries there

```{r}
tb_covid %>% head(5)
tb_demo %>% head(5) #%>% show_query()

```


#### Bit more complex queries

1. Get the latest data for each county
2. Get the data on 1 Aug, 2020, and select county with more than 10 deaths

```{r}
tb_covid %>%
  filter(date == '2021-03-19') %>%
  arrange(desc(deaths)) 

```
#### Ordering and aggregation

1. Order counties with the highest number of death
2. Calculate the state level death figures


```{r}
tb_covid %>%
  filter(date == '2021-03-19') %>% 
  arrange(desc(deaths))
tb_covid %>%
  filter(date == '2021-03-19') %>% 
  group_by(state) %>% summarize(tot_deaths = sum(deaths))

```


### Merging tables

- merge covid and demo table by fips
- calculate per 1000 cases and deaths


```{r}
tb_covid %>% 
  left_join(tb_demo, by = "fips") %>% 
  mutate(deaths_per1000 = deaths/TotalPop *1000) # I thought this line is not working, but actually it is...

```